---
title: "Solar Panel Business Case"
date: '`r Sys.Date()`'
output: 
  html_document:
    includes:
      in_header: header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

require(ggplot2)
require(stringr)
require(lubridate)
```

```{r data, include=FALSE}
stopifnot(file.exists("ActualData.csv"))
stopifnot(file.exists("ForecastData.csv"))

actual <- read.csv("ActualData.csv")
names(actual)[1] <- "Bill.Start"

strip_dollar_sign_or_comma <- function(x) as.numeric(gsub("[,$]", "", x))

forecast <- read.csv("ForecastData.csv")
names(forecast)[1] <- "Year"

forecast$yearmonth <- paste0(forecast$Year, "-", stringr::str_pad(forecast$Month, pad = "0", width = 2))
forecast$kWh <- strip_dollar_sign_or_comma(forecast$kWh)
forecast$Transmission <- strip_dollar_sign_or_comma(forecast$Transmission)
forecast$Customer <- strip_dollar_sign_or_comma(forecast$Customer)
forecast$Distribution <- strip_dollar_sign_or_comma(forecast$Distribution)
forecast$Generation <- strip_dollar_sign_or_comma(forecast$Generation)
forecast$Bill.Total <- strip_dollar_sign_or_comma(forecast$Bill.Total)

actual$Bill.Start <- lubridate::mdy(actual$Bill.Start)
actual$Bill.End <- lubridate::mdy(actual$Bill.End)

# validate Dates
stopifnot(all(as.numeric(with(actual, Bill.End - Bill.Start - Days + 1)) == 0))
# validate Year
stopifnot(all(actual$Year - lubridate::year(actual$Bill.End) == 0))
# validate Month
stopifnot(all(actual$Month - lubridate::month(actual$Month) == 0))
```

## Introduction

## Pre-Installation Forecast

Goal: Determine which, if any, of the solar quotes have a positive net present
value based on historical usage data.

### Usage Data (Apr 2019 to July 2022)

```{r}
ggplot(forecast, aes(x = yearmonth, y = kWh)) + geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  geom_vline(aes(xintercept = "2020-03"), col = "red", lty = 2) +
  geom_text(x = "2020-03", y = 1200, label = "EV Purchased", angle = 90, vjust = 1) +
  labs(x = "", y = "kWh")
```

```{r include=FALSE}
actual_usage_previous_year <- sum(forecast$kWh[(nrow(forecast)-11):nrow(forecast)])
actual_cost_previous_year <- sum(forecast$Bill.Total[(nrow(forecast)-11):nrow(forecast)])
forecast_usage_per_year <- mean(forecast$kWh[16:length(forecast$kWh)])*12
forecast_trans_per_year <- mean(forecast$Transmission[28:40])*12
forecast_dist_per_year <- mean(forecast$Distribution[28:40])*12
forecast_cust_per_year <- forecast$Customer[40]*12
forecast_gen_per_year <- mean(forecast$kWh[16:length(forecast$kWh)])*12 * forecast$Generation.per.kWh[40]
forecast_bill_total <- forecast_trans_per_year + forecast_dist_per_year + forecast_cust_per_year + forecast_gen_per_year
blue_raven_production <- 13380
ecohouse_production <- 15453
blue_raven_eff <- .92
ecohouse_eff <- 0.8
blue_raven_total_cost <- 46341
blue_raven_tax_break <- 12049
eco_total_cost <- 42500
eco_tax_break <- 12750
energy_inflation <- 0.04
inflation <- 0.02
```

- Usage per year (KWh): `r scales::comma(forecast_usage_per_year)`
- Transmission Cost per year: `r scales::dollar(forecast_trans_per_year)`
- Distribution Cost per year: `r scales::dollar(forecast_dist_per_year)`
- Customer Cost per year: `r scales::dollar(forecast_cust_per_year)`
- Generation Cost per year: `r scales::dollar(forecast_gen_per_year)`
- Bill Total: `r scales::dollar(forecast_bill_total)`

### Options for Solar Systems

- Blue Raven:
    - Panels: 32
    - W per panel: 400
    - Total Generation (KW): 12.8 = 32*400
    - Yearly (KWh): 13,380
    - Total Cost: `r scales::dollar(blue_raven_total_cost)`
    - Tax Break: `r scales::dollar(blue_raven_tax_break)`
    - Net Cost: `r scales::dollar(blue_raven_total_cost-blue_raven_tax_break)`
    - 25 yr Efficiency: `r scales::percent(blue_raven_eff)`
- EcoHouse
    - Panels: 39
    - W per panel: 360
    - Total Generation (KW): 14.04 = 39*360
    - Yearly (KWh): 15,453
    - Total Cost: `r scales::dollar(eco_total_cost)`
    - Tax Break: `r scales::dollar(eco_tax_break)`
    - Net Cost: `r scales::dollar(eco_total_cost-eco_tax_break)`
    - 25 yr Efficiency: `r scales::percent(ecohouse_eff)`

### Assumptions:

- Energy Inflation: `r scales::percent(energy_inflation)`
- Overall Inflation: `r scales::percent(inflation)`

```{r}
npv_data <- data.frame(
  year = 2022:2052,
  usage = c(actual_usage_previous_year, rep(forecast_usage_per_year, 30)),
  cost_no_solar = c(actual_cost_previous_year, 
                   forecast_bill_total*(1+energy_inflation)^(1:30)),
  generation_blue = c(0, blue_raven_production*(blue_raven_eff^(1/25))^(1:30)),
  generation_eco = c(0, ecohouse_production*(ecohouse_eff^(1/25))^(1:30))) |>
  dplyr::mutate(
    offset_blue = c(-(blue_raven_total_cost-blue_raven_tax_break), rep(0, 30)) +
      generation_blue*forecast$Generation.per.kWh[40]*(1+energy_inflation)^(0:30),
    offset_eco = c(-(eco_total_cost-eco_tax_break), rep(0, 30)) +
      generation_eco*forecast$Generation.per.kWh[40]*(1+energy_inflation)^(0:30))
```

### Forecast Timeseries

```{r}
ggplot(npv_data, aes(x = year, y = cumsum(usage))) + 
  geom_point() + 
  geom_line() +
  labs(x = "", y = "Cumulative Energy Usage (kWh)") +
  scale_y_continuous(labels = scales::comma_format())
```

```{r}
ggplot(npv_data |>
         reshape2::melt(id.vars = "year") |>
         dplyr::filter(variable %in% c("cost_no_solar", "offset_blue", "offset_eco")) |>
         dplyr::group_by(variable) |>
         dplyr::mutate(value = cumsum(value)) |>
         dplyr::ungroup() |>
         dplyr::mutate(variable = factor(variable, labels = c("Cumulative Base Cost", "Cumulative value of Blue Raven", "Cumulative value of Ecohouse"))), 
       aes(x = year, y = value, group = variable, col = variable)) + 
  geom_point() + 
  geom_line() +
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(x = "", y = "", col = "")
```

### Calculate Forecasted 25 yr Net Present Value

- Blue Raven: `r scales::dollar(sum(npv_data$offset_blue[1:26]*(1-inflation)^(0:25)))`
- EcoHouse: `r scales::dollar(sum(npv_data$offset_eco[1:26]*(1-inflation)^(0:25)))`

The Ecohouse solar options was selected

## Pre-Installation Temperature vs Energy Forecast

```{r, include=FALSE}
model_data <- data.frame(kWh = forecast$kWh[-1],
                         AvgTemp = forecast$AvgTemp[-1],
                         AvgTemp2 = forecast$AvgTemp[-1]^2,
                         Days = forecast$Days[-1],
                         kWh_per_day = forecast$kWh[-1] / forecast$Days[-1],
                         month = forecast$Month[-1])
lm1 <- lm(kWh_per_day ~ AvgTemp + AvgTemp2, data = model_data)

model_data$pred <- predict(lm1)
```

### Monthly Temperature Timeseries

```{r}
ggplot(forecast, aes(x = yearmonth, y = AvgTemp)) +
  geom_point() +
  labs(x = "Bill Month", y = "Average Temperature (F)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

### Energy Usage vs Temperature

```{r}
ggplot(model_data, aes(x = AvgTemp, y = kWh)) +
  geom_point() +
  labs(x = "Month Average Temperature", y = "Monthly Total kWh")
```

### Energy Usage per Day vs Temperature

Energy usage per day should remove some variability due to months of different lengths

```{r}
ggplot(model_data, aes(x = AvgTemp, y = kWh_per_day)) +
  geom_point() +
  geom_line(aes(x = AvgTemp, y = pred), col = "red", lty = 2) +
  labs(x = "Month Average Temperature", y = "kWh Per Day")
```

### Regression Diagnositics and Fit

```{r}
plot(lm1, which = 1)
plot(lm1, which = 2)
```

## Actuals and Forecast (Post-Installation)

```{r}
X <- actual |>
  dplyr::filter(Bill.Start > as.Date("2022-08-01")) |>
  dplyr::rename(total_kwh = kWh,
                delivered_kwh = Delivered..kWh.,
                received_kwh = Received..kWh.,
                generation_supply = Generation..Supply.) |>
  dplyr::mutate(home_used = delivered_kwh + received_kwh + Solar.Production,
                transmission_per_kwh = Transmission / total_kwh,
                distribution_per_kwh = Distribution / total_kwh,
                forecast_distribution_per_kwh = ifelse(total_kwh > 0, Distribution / total_kwh, 0))

for (i in 2:nrow(X)) {
  if (X$transmission_per_kwh[i] == 0) X$transmission_per_kwh[i] <- X$transmission_per_kwh[i-1]
  if (X$forecast_distribution_per_kwh[i] == 0) X$forecast_distribution_per_kwh[i] <- X$forecast_distribution_per_kwh[i-1]
}

Y <- data.frame(Bill.Start = rep(as.Date(NA), 400),
                Bill.End = rep(as.character(NA), 400),
                Year = rep(NA_integer_, 400),
                Month = rep(NA_integer_, 400),
                Days = rep(NA_integer_, 400),
                total_kwh = rep(NA_integer_, 400),
                delivered_kwh = rep(NA_integer_, 400),
                received_kwh = rep(NA_integer_, 400),
                Transmission = rep(as.numeric(NA), 400),
                generation_supply = rep(as.numeric(NA), 400),
                Distribution = rep(as.numeric(NA), 400),
                Customer = rep(NA_integer_, 400),
                Generation.per.kWh = rep(as.numeric(NA), 400),
                Generation = rep(as.numeric(NA), 400),
                AvgTemp = rep(NA_integer_, 400),
                Solar.Production = rep(as.numeric(NA), 400),
                Bill.Total = rep(as.numeric(NA), 400),
                home_used = rep(as.numeric(NA), 400),
                transmission_per_kwh = rep(as.numeric(NA, 400)),
                distribution_per_kwh = rep(as.numeric(NA, 400)),
                forecast_distribution_per_kwh = rep(as.numeric(NA, 400)))
X2 <- rbind(X, Y)

for (i in (nrow(X) + 1):(nrow(X) + 400)) {
  X2$Month[i] = X2$Month[i-12]
  X2$Year[i] = X2$Year[i-12] + 1
  ind <- which(X$Month == X2$Month[i])
  X2$Days[i] = mean(X$Days[ind])
  X2$AvgTemp[i] = mean(X$AvgTemp[ind])
  X2$transmission_per_kwh[i] = X2$transmission_per_kwh[i-1]
  X2$forecast_distribution_per_kwh[i] = X2$forecast_distribution_per_kwh[i-1]
  X2$Generation.per.kWh[i] = X2$Generation.per.kWh[i-1]
}

X2$forecast_total_kwh = pmax(X2$home_used, (coef(lm1)[1] + coef(lm1)[2]*X2$AvgTemp + coef(lm1)[3]*X2$AvgTemp^2)*X2$Days, na.rm = TRUE)

X2 <- X2 |>
  dplyr::mutate(forecast_transmission = forecast_total_kwh * transmission_per_kwh,
                forecast_distribution = forecast_total_kwh * forecast_distribution_per_kwh,
                forecast_customer = 10,
                forecast_generation = forecast_total_kwh * Generation.per.kWh,
                forecast_total = forecast_transmission + forecast_distribution + forecast_customer + forecast_generation,
                estimated_save = ifelse(!is.na(Bill.Total), forecast_total - Bill.Total, forecast_total),
                total_cash_flow = estimated_save)

X2$total_cash_flow[1] <- X2$total_cash_flow[1] - 42500
X2$total_cash_flow[7] <- X2$total_cash_flow[7] + 12750

X2$npv <- 0
i_switch <- 0
X2$npv[1] <- X2$total_cash_flow[1]
for (i in 2:nrow(X2)) {
  X2$npv[i] <- sum(X2$total_cash_flow[1:i]*(1-0.02/12)^(0:(i-1)))
  if (X2$npv[i] >= 0 & X2$npv[i-1] < 0) i_switch = i
}

X2$year_month <- paste0(X2$Year, "-", stringr::str_pad(X2$Month, pad = "0", width = 2))
```

#### Actual Temp

```{r}
ggplot(X2 |>
         dplyr::filter(!is.na(Bill.Start)), aes(x = Bill.Start, y = AvgTemp)) + 
  geom_point() +
  labs(x = "", y = "Month Average Temperature (F)")
```

#### Actual Usage since Installation

- **Total Demand** - The amount of power used in the home
- **From Grid** - The amount of power needed by the house from the grid
- **To Grid** - The amount of power supplied by the solar panels to the grid
- **Solar Production** - The amount of power created by the solar panels

$$Solar\ Production\ +\ From\ Grid = Actual\ Total\ Demand\ -\ To\ Grid$$

$$To\ Grid < 0$$


```{r}
ggplot(X2 |>
         dplyr::filter(!is.na(Bill.Start)) |>
         dplyr::select(Bill.Start, forecast_total_kwh, delivered_kwh, received_kwh, Solar.Production) |>
         reshape2::melt(id.vars = "Bill.Start") |>
         dplyr::mutate(variable = factor(variable, labels = c("Actual Total Demand",
                                                              "From Grid",
                                                              "To Grid",
                                                              "Solar Production"))), 
       aes(x = Bill.Start, y = value, group = variable, col = variable)) + 
  geom_point() +
  labs(x = "", y = "Power (kWh)", col = "") +
  scale_y_continuous(labels = scales::comma_format()) +
  geom_vline(xintercept = as.Date("2025-01-31"), lty = 2) +
  geom_text(x = as.Date("2025-03-01"), y = 0, label = "Mach-E EV Purchased", angle = 90, vjust = 1, col = "black", size = 3)
```

### Comparison of Demand and Solar Production

```{r}
ggplot(X2 |>
         dplyr::filter(!is.na(Bill.Start)) |>
         dplyr::select(Bill.Start, forecast_total_kwh, Solar.Production) |>
         reshape2::melt(id.vars = "Bill.Start") |>
         dplyr::mutate(variable = factor(variable, labels = c("Actual Total Demand",
                                                              "Solar Production"))), 
       aes(x = Bill.Start, y = value, group = variable, col = variable)) + 
  geom_point() +
  geom_line() +
  labs(x = "", y = "Power (kWh)", col = "") +
  scale_y_continuous(labels = scales::comma_format()) +
  geom_vline(xintercept = as.Date("2025-01-31"), lty = 2) +
  geom_text(x = as.Date("2025-03-01"), y = 2000, label = "Mach-E EV Purchased", angle = 90, vjust = 1, col = "black", size = 3)

```

### Actual Demand vs Predicted Demand from Avg Temp

```{r warning=FALSE}
temp <- X2 |>
  dplyr::filter(!is.na(Bill.End)) |>
  dplyr::mutate(
    predicted = predict(lm1, newdata = data.frame(AvgTemp = AvgTemp,
                                  AvgTemp2 = AvgTemp^2)) * Days,
    actual = forecast_total_kwh
  )

ggplot(temp, aes(x = actual, y = predicted)) + 
  geom_point() +
  labs(x = "Actual Total Demand", y = "Predicted Total Demand") +
  geom_abline(slope = 1, intercept = 0, lty = 2, col = "darkgrey") +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma)

ggplot(temp |>
         dplyr::select(Bill.End, predicted, actual) |>
         reshape2::melt(id.vars = "Bill.End"),
       aes(x = Bill.End, y = value, col = variable, group = variable)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "", y = "Total Demand", col = "")

```

```{r}
temp <- X2 |>
  dplyr::mutate(forecast_total_kwh = cumsum(forecast_total_kwh)) |>
  dplyr::filter(Month == 12)
ggplot(temp, aes(x = Year, y = forecast_total_kwh)) + 
  geom_point() +
  geom_line() +
  labs(x = "", y = "Forecasted Cumulative Total kWh since Installation") +
  scale_y_continuous(labels = scales::comma_format())
```

```{r}
temp <- X2 |>
  dplyr::mutate(cum_total_cash_flow = cumsum(total_cash_flow)) |>
  dplyr::filter(Month == 12, Year <= 2052) |>
  dplyr::left_join(npv_data |>
                     dplyr::select(year, offset_eco),
                   by = c("Year" = "year")) |>
  dplyr::mutate(cum_orig_cash_flow = cumsum(offset_eco))
temp2 <- X2 |>
  dplyr::group_by(Year) |>
  dplyr::summarize(total_cash_flow = sum(total_cash_flow)) |>
  dplyr::filter(Year <= 2052)
ggplot(temp, aes(x = Year, y = cum_total_cash_flow)) + 
  geom_point() +
  geom_line() +
  geom_point(mapping = aes(x = Year, y = cum_orig_cash_flow), col = "green", shape = 1) +
  geom_line(mapping = aes(x = Year, y = cum_orig_cash_flow), col = "green", lty = 2) +
  labs(x = "", y = "Forecasted Nominal Cumulative Cash Flow") +
  geom_hline(yintercept = 0, lty = 2, col = "red") +
  scale_y_continuous(labels = scales::dollar_format()) +
  geom_col(data = temp2, mapping = aes(x = Year, y = total_cash_flow), alpha = 0.5)
```

```{r}
ggplot(temp, aes(x = Year, y = npv)) + 
  geom_point() +
  geom_line() +
  labs(x = "", y = "Forecasted Real Cumulative Cash Flow\n2022 Dollars") +
  geom_hline(yintercept = 0, lty = 2, col = "red") +
  geom_vline(xintercept = X2$Year[i_switch], lty = 2, col = "red") +
  scale_y_continuous(labels = scales::dollar_format())

```

### Forecast Return on Investment

- Break Even Year:  `r X2$Year[i_switch]`
- 20 yr NPV in 2042: `r ind <- which(X2$Year == 2042 & X2$Month == 8); scales::dollar(X2$npv[ind])`
- 25 yr NPV in 2047: `r ind <- which(X2$Year == 2047 & X2$Month == 8); scales::dollar(X2$npv[ind])`

